### 2020-03-06 short.body.tex
Working.

Epidemic outbreaks include a spectrum of illness, including
asymptomatic cases that can be mild and/or subclinical.
As such, symptomatic cases are the predominant focus of treatment and usually represent
the bulk of reported cases. However, infected individuals who are asymptomatic
yet infectious can be a critical, albeit poorly characterized
factor in the spread of pathogens, as well as 
present challenges for control.
Asymptomatic individuals
are hard to trace, unlikely to self-isolate,
may move outside of quarantine,  
retain normal social and travel patterns, and potentially spark new
outbreaks.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
%For example, the incidence of asymptomatic
%infection was estimated to be
%approximately 33\% (or higher) 
%from an outbreak of COVID-19 on the Princess Cruises ship~\citep{mizumoto_2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?
%.(i) What fraction of newly infected individuals
%have mild and/or subclinical infections rather than severe and/or
%clinical infections? (ii) What is the time-scale over which 
%individuals who have mild and/or subclinical cases remain
%infectious? (iii) How infectious are individuals
%with mild and/or subclinical infections relative to those
%with severe symptoms?  
Here, we attempt to address the impact of asymptomatic cases differently, by relating 
individual-level features of asymptomatic cases (e.g., the probability
of initiating an asymptomatic case, asymptomatic case duration, and
the transmission rates from asymptomatic individuals) to dynamics at the
population scale. That is, rather than focusing on the prevalence
of asymptomatic cases, we are interested in understanding the
relevance of asymptomatic infections to secondary case production -- including
to the production of symptomatic cases.
%Our purpose is to characterize key drivers of contribution to
%secondary cases during an outbreak by `asymptomatic' individuals --
%and help characterize the relevance of asymptomatic transmission to
%outbreak dynamics and control.

To begin, consider a renewal equation variant in which incidence $i$ is divided into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively. 
Both types of individuals can infect others, but, they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$, reflecting their differences in the abilities to transmit and the courses of infection.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
an individual with subclinical, and fast-clearing, symptoms is likely to have short generation intervals, whereas an individual with mild, but long-lasting, symptoms is likely to have long generation intervals.
In total, the dynamics of susceptibles and incidence are:
\begin{eqnarray}
\dot{S}&=&-i(t) \\
i(t)&=&\mathcal R_a S \int_0^\infty i_a(t-\tau) g_a(\tau) \mathrm{d}\tau + \mathcal R_s S \int_0^\infty i_s(t-\tau) g_s(\tau) \mathrm{d}\tau.
\end{eqnarray}
Here, $p$ is the fraction of asymptomatic cases that are generated for each infected individual ($i_a(t)= p i(t)$), and $1-p$ is the fraction of symptomatic cases that are generated for each infected individual ($i_s(t)= (1-p) i(t)$).

%To begin, consider a SEIR model variant in which there are two infectious categories --
%$I_a$ and $I_s$ -- corresponding to asymptomatic and symptomatic cases,
%respectively. Both types of individuals can infect others, but,
%only symptomatic cases can lead to fatalities, denoted
%by the $D$ category. In total, the dynamics of susceptibles, exposed,
%infectious, recovered, and dead are:
%\begin{eqnarray}
%\dot{S}&=&-\beta_a S I_a -\beta_s S I_s \\
%\dot{E}&=&\beta_a S I_a +\beta_s S I_s -\gamma_e E\\
%\dot{I}_a&=&p\gamma_e E-\gamma_a I_a\\
%\dot{I}_s&=&(1-p)\gamma_e E-\gamma_s I_s\\
%\dot{R}&=&\gamma_a I_a + (1-f)\gamma_s I_s \\
%\dot{D}&=&f\gamma_s I_s.
%\end{eqnarray}
%Here, $\beta_a$ and $\beta_s$ denote transmission rates,
%$\gamma_e$ denotes the transition from exposed to infectious,
%$p$ is the fraction of asymptomatic cases that
%are generated for each exposed individual,
%$1-p$ is the fraction of symptomatic cases that
%are generated for each exposed individual,
%$\gamma_a$ and $\gamma_s$ denote recovery rates,
%and $f$ denotes the case fatality ratio for symptomatic cases.

The basic reproduction number, ${\cal{R}}_0$, characterize the strength of an epidemic in an otherwise susceptible population \citep{anderson1992infectious}:
\begin{equation}
{\cal{R}}_{0}= p {\cal{R}}_a + (1-p) {\cal{R}}_s.
\end{equation}
We define the intrinsic proportions of asymptomatic and symptomatic transmission as their relative contribution toward the basic reproduction number:
\begin{eqnarray}
z &=& p {\cal{R}}_a/{\cal{R}}_{0},\\
1-z &=& (1-p) {\cal{R}}_s/{\cal{R}}_{0}.
\end{eqnarray}
Then, the intrinsic generation-interval distribution depends on the intrinsic proportion of asymptomatic vs.~symptomatic transmission and their corresponding intrinsic generation-interval distributions:
\begin{equation}
g(\tau) = z g_a(\tau) + (1-z) g_s(\tau),
\end{equation}
which allows us to express incidence of infection as a renewal process that depends on previous incidence:
\begin{equation}
i(t) = {\cal{R}}_{0} \int_0^\infty i(t-\tau) g(\tau) \mathrm{d} \tau.
\end{equation}
Yet, this information is not sufficient to disentangle the role of asymptomatic cases, i.e., what fraction of secondary cases can be ascribed to \emph{realized} transmission from asymptomatic cases vs.~symptomatic cases?

>>>>

Epidemic outbreaks include a spectrum of illness, including
asymptomatic cases that can be mild and/or subclinical.
As such, symptomatic cases are the predominant focus of treatment and usually represent
the bulk of reported cases. However, infected individuals who are asymptomatic
yet infectious can be a critical, albeit poorly characterized
factor in the spread of pathogens, as well as 
present challenges for control.
Asymptomatic individuals
are hard to trace, unlikely to self-isolate,
may move outside of quarantine,  
retain normal social and travel patterns, and potentially spark new
outbreaks.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
%For example, the incidence of asymptomatic
%infection was estimated to be
%approximately 33\% (or higher) 
%from an outbreak of COVID-19 on the Princess Cruises ship~\citep{mizumoto_2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?
%.(i) What fraction of newly infected individuals
%have mild and/or subclinical infections rather than severe and/or
%clinical infections? (ii) What is the time-scale over which 
%individuals who have mild and/or subclinical cases remain
%infectious? (iii) How infectious are individuals
%with mild and/or subclinical infections relative to those
%with severe symptoms?  
There is currently a lot of interest in the role of asymptomatic infections in COVIF for two major reasons. 
First, the proportion of infections that are asymptomatic is critical to attempts to estimate the likely burden of severe outcomes (including mortality) when the virus spreads through a population.
Second, understanding the possible role of \emph{transmission} by asymptomatic individuals is crucial to planning surveillance and control efforts.

Here, we focus on a third effect. 
If asymptomatic cases are important for transmission, they also have the potential to affect estimates of key epidemiological parameters. 
Here, we investigate the relationship between individual-level features of asymptomatic cases (e.g., the probability of initiating an asymptomatic case, asymptomatic case duration, and the transmission rates from asymptomatic individuals) to dynamics at the population scale. 
That is, rather than focusing on the prevalence
of asymptomatic cases, we are interested in understanding the
relevance of asymptomatic infections to secondary case production -- including
to the production of symptomatic cases.
%Our purpose is to characterize key drivers of contribution to
%secondary cases during an outbreak by `asymptomatic' individuals --
%and help characterize the relevance of asymptomatic transmission to
%outbreak dynamics and control.

We model viral spread using a renewal-equation framework, with incidence $i$ is divided into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively. 
Both types of individuals can infect others, but they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
individuals with subclinical cases may have fast clearance and short generation intervals, or slow viral reproduction and long generation intervals.

Neglecting births and loss of immunity on the time scale of the outbreak, the dynamics of susceptibles and incidence are:
\begin{eqnarray}
\dot{S}&=&-i(t) \\
i(t)&=&\mathcal R_a S \int_0^\infty i_a(t-\tau) g_a(\tau) \mathrm{d}\tau + \mathcal R_s S \int_0^\infty i_s(t-\tau) g_s(\tau) \mathrm{d}\tau.
\end{eqnarray}
Here, $p$ is the proportion of \emph{incident cases} that are asymptomatic.

The basic reproduction number of this system \citep{anderson1992infectious} is:
\begin{equation}
{\cal{R}}_{0}= p {\cal{R}}_a + (1-p) {\cal{R}}_s.
\end{equation}
We define the ``intrinsic'' proportion of asymptomatic transmission as the relative contribution of asymptomatic cases to the basic reproduction number:
\begin{equation}
z = p {\cal{R}}_a/{\cal{R}}_{0}
\end{equation}
Note that the intrinsic proportion of symptomatic transmission satisfies
\begin{equation}
1-z &=& (1-p) {\cal{R}}_s/{\cal{R}}_{0}.
\end{equation}

Then, the intrinsic generation-interval distribution depends on the intrinsic proportion of asymptomatic vs.~symptomatic transmission and their corresponding intrinsic generation-interval distributions:
\begin{equation}
g(\tau) = z g_a(\tau) + (1-z) g_s(\tau),
\end{equation}
which allows us to express incidence of infection as a renewal process that depends on previous incidence:
\begin{equation}
i(t) = {\cal{R}}_{0} \int_0^\infty i(t-\tau) g(\tau) \mathrm{d} \tau.
\end{equation}
Yet, this information is not sufficient to disentangle the role of asymptomatic cases, i.e., what fraction of secondary cases can be ascribed to \emph{realized} transmission from asymptomatic cases vs.~symptomatic cases?
